\documentclass[12pt]{article}

% =========================
% USPTO-SAFE FORMATTING
% =========================
\usepackage[letterpaper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}

% USPTO spacing: 1.5 or double
\usepackage{setspace}
\setstretch{1.5}

% Paragraph control
\usepackage{parskip}

% Math
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}

% Lists
\usepackage{enumitem}
\setlist[itemize]{nosep}
\setlist[enumerate]{nosep}

% Links (safe breaking)
\usepackage[hidelinks]{hyperref}
\usepackage{xurl} % allows line breaks in URLs if any appear

% Figures placeholders
\usepackage{graphicx}
\usepackage{caption}

% =========================
% OVERFLOW / LINE-BREAK CONTROL (IMPORTANT)
% =========================
\setlength{\emergencystretch}{2em} % gives TeX extra flexibility before overfull boxes
\sloppy                             % relaxes line breaking (reduces overfull hboxes)
\microtypesetup{expansion=true,protrusion=true}

% If you ever have long, unbreakable identifiers, these help:
\hyphenation{meta-cog-ni-tive ge-ne-ra-tive se-man-tic en-tro-py fed-er-at-ed}

% =========================
% Equation numbering (sequential, global)
% =========================
\renewcommand{\theequation}{\arabic{equation}}

% =========================
% Paragraph numbering [0001], [0002], ...
% =========================
\newcounter{para}
\setcounter{para}{0}
\newcommand{\p}{%
  \refstepcounter{para}%
  \noindent[\ifnum\value{para}<10 000\fi
            \ifnum\value{para}<100 0\fi
            \ifnum\value{para}<1000 0\fi
            \arabic{para}] \ignorespaces}

% =========================
% Simple macros
% =========================
\newcommand{\GenAI}{GenAI}
\newcommand{\BOS}{BOS}
\newcommand{\FDM}{FDM}
\newcommand{\MIA}{MIA}
\newcommand{\MVL}{MVL}
\newcommand{\SEP}{SEP}
\newcommand{\EKF}{EKF}

\begin{document}

\begin{center}
{\Large \textbf{SPECIFICATION}}\\[4pt]
\end{center}

\section*{TITLE OF THE INVENTION}
\p \textbf{Behavioral Orchestration System for Regulating Learner State in Generative AI Learning Environments}

\section*{CROSS-REFERENCE TO RELATED APPLICATIONS}
\p Not Applicable.

\section*{STATEMENT REGARDING FEDERALLY SPONSORED RESEARCH OR DEVELOPMENT}
\p Not Applicable.

\section*{BACKGROUND OF THE INVENTION}

\subsection*{Field of the Invention}
\p The invention relates to educational technology and generative artificial intelligence, and in particular to computer-implemented systems and methods for closed-loop regulation of learner interaction, integrity scaffolding, and orchestration of cognitive, affective, and metacognitive states in AI-mediated learning environments.

\subsection*{Description of Related Art}
\p Existing intelligent tutoring systems often optimize content sequencing based primarily on cognitive mastery estimates. Conversational generative-AI tutors introduce additional risks including unreliable outputs, cognitive offloading, and reduced learner autonomy. Post-hoc integrity detection can be brittle and does not directly improve the interactive computing loop. There remains a need for a technological solution that improves the functioning and stability of real-time, computer-implemented generative-AI learning systems while preserving privacy and teacher agency.

\section*{BRIEF SUMMARY OF THE INVENTION}

\p \textbf{Behavioral Orchestration System (\BOS).}
The \BOS\ senses interaction signals, estimates a latent learner state, and applies interventions to regulate a real-time learning interaction loop toward a target region. In embodiments, the \BOS\ comprises a Feature Detection Module (\FDM), a State Estimator (including an \EKF), an Orchestrator (control policy), and a Metacognitive Integrity Architecture (\MIA) that can trigger a Metacognitive Verification Loop (\MVL).

\p \textbf{Technological problem and technological solution.}
The \BOS\ addresses technical problems of: (i) unstable real-time interactive control in the presence of noisy and sparse signals, (ii) high computational latency and processing cost of uncertainty and integrity estimation for generative-AI outputs, and (iii) privacy and communication inefficiency caused by transmitting raw interaction data. The \BOS\ provides corresponding technical solutions including \EKF-based state estimation for control-loop stability, semantic-entropy probes for computational efficiency, and privacy-by-design via on-device feature extraction and federated learning to improve data security and network efficiency.

\section*{BRIEF DESCRIPTION OF THE SEVERAL VIEWS OF THE DRAWINGS}
\p The following drawings are described as non-limiting examples:
\begin{itemize}
  \item \textbf{FIG. 1} is an overall architecture of the \BOS\ integrated with a generative-AI tutor, local feature extraction, and teacher override.
  \item \textbf{FIG. 2} illustrates a state-space model of latent learner state, control inputs, and observations.
  \item \textbf{FIG. 3} illustrates an \EKF\ prediction and measurement update flow with missing-observation handling.
  \item \textbf{FIG. 4} illustrates an \MVL\ gating flow driven by integrity risk and corroboration constraints.
  \item \textbf{FIG. 5} illustrates semantic-entropy estimation via sampling and an \SEP\ approximation path from single-generation internal signals.
  \item \textbf{FIG. 6} illustrates privacy-by-design via on-device feature extraction and federated learning with aggregation.
  \item \textbf{FIG. 7} illustrates a teacher override/blending mechanism, audit logging, and intervention rate limiting.
\end{itemize}

\section*{DETAILED DESCRIPTION OF THE INVENTION}

\subsection*{0. Definitions and Notation (Non-Limiting)}
\p Time is indexed by $t \in \{0,1,2,\dots\}$. A latent learner state vector is denoted $\bm{x}_t \in \mathbb{R}^n$, an observation vector is denoted $\bm{y}_t \in \mathbb{R}^m$, and an intervention/control vector is denoted $\bm{u}_t \in \mathbb{R}^p$.

\p \textbf{Module terminology (structural mitigation).}
The term ``module'' as used herein refers to specific structure implemented in hardware and/or software, including: one or more processors executing stored instructions from non-transitory memory, an ASIC, an FPGA, and/or a specialized microprocessor or accelerator executing instructions, alone or in combination.

% (… keep the rest of your spec exactly as previously provided …)

\newpage
\section*{CLAIMS}

\begin{enumerate}[label=\arabic*., leftmargin=*, itemsep=6pt]
  \item A computer-implemented educational orchestration system comprising:
  \begin{enumerate}[label=(\alph*), leftmargin=*, itemsep=4pt]
    \item a feature detection module configured to obtain interaction data associated with a learner interacting with a generative artificial intelligence system and to generate an observation vector from the interaction data;
    \item a state estimator comprising an extended Kalman filter configured to estimate a latent learner state vector comprising at least a cognitive component, an affective component, and a metacognitive component, based on the observation vector and a state-space model;
    \item a metacognitive integrity architecture configured to approximate a hallucination risk using a semantic entropy probe derived from internal signals of a single generation to reduce computational latency and processing cost; and
    \item an orchestrator configured to select an intervention vector to be applied to the learner interaction to regulate the latent learner state vector toward a target state region,
  \end{enumerate}
  wherein the orchestrator selects the intervention vector based at least in part on minimizing a cost function that includes an autonomy cost penalizing over-intervention.

  \item A computer-implemented method of regulating learner interaction in a generative artificial intelligence learning environment, the method comprising:
  \begin{enumerate}[label=(\alph*), leftmargin=*, itemsep=4pt]
    \item obtaining interaction data from learner interaction with a generative artificial intelligence system;
    \item generating an observation vector from the interaction data;
    \item estimating, using a state estimator comprising an extended Kalman filter, a latent learner state vector comprising cognitive, affective, and metacognitive components based on the observation vector and a state-space model;
    \item approximating, using a metacognitive integrity architecture, a hallucination risk using a semantic entropy probe derived from internal signals of a single generation to reduce computational latency and processing cost;
    \item selecting an intervention vector based on a control policy minimizing a cost function including an autonomy cost; and
    \item applying the intervention vector to modify the learner interaction in the learning environment.
  \end{enumerate}
\end{enumerate}

\newpage
\section*{ABSTRACT OF THE DISCLOSURE}
\noindent
A computer-implemented system regulates learner interaction in a generative artificial intelligence learning environment using closed-loop control. Interaction signals are transformed into an observation vector and used to estimate a latent learner state comprising cognitive, affective, and metacognitive components using a stochastic state-space model and an extended Kalman filter. A metacognitive integrity architecture approximates hallucination risk using a semantic entropy probe derived from internal signals of a single generation to reduce latency and processing cost, and triggers non-punitive verification actions when integrity risk is detected. An orchestrator selects interventions to maintain the learner state within a target region while minimizing a cost function including an autonomy cost. Privacy-preserving operation may be provided by on-device feature extraction and federated learning that transmits parameter updates rather than raw data, with teacher agency supported via override and blending.

\end{document}